---
title: "activity02"
output: html_document
date: "`r Sys.Date`"
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(ggplot2)
```

### The data

The data we're working with is from the OpenIntro site: `https://www.openintro.org/data/csv/hfi.csv`.
Here is the "about" page: <https://www.openintro.org/data/index.php?data=hfi>.

In the R code chunk below titled `load-data`, you will type the code that reads in the above linked CSV file by doing the following:

-   Rather than downloading this file, uploading to RStudio, then reading it in, explore how to load this file directly from the provided URL with `readr::read_csv` (`{readr}` is part of `{tidyverse}`).
-   Assign this data set into a data frame named `hfi` (short for "Human Freedom Index").

```{r load data}
hfi <- read_csv("hfi.csv")
hfi
```

1.  What are the dimensions of the dataset? What does each row represent?

```{r}
dimensions <- dim(hfi)
rows <- dimensions[1]
columns <- dimensions[2]

# Print the dimensions
cat("The dataset has", rows, "rows and", columns, "columns.\n")
```

The dataset spans a lot of years.
We are only interested in data from year 2016.
In the R code chunk below titled `hfi-2016`, type the code that does the following:

-   Filter the data `hfi` data frame for year 2016, and
-   Assigns the result to a data frame named `hfi_2016`.

```{r hfi-2016}
library(dplyr)

# Filter the data for year 2016 and assign it to hfi_2016
hfi_2016 <- hfi |>
  filter(year == 2016)
hfi_2016
```

### 1. Identify our research question(s)

The research question is often defined by you (or your company, boss, etc.).
Today's research question/goal is to predict a country's personal freedom score in 2016.

For this activity we want to explore the relationship between the personal freedom score, `pf_score`, and the political pressures and controls on media content index,`pf_expression_control`.
Specifically, we are going to use the political pressures and controls on media content index to predict a country's personal freedom score in 2016.

### 2. Explore the variables of interest

Answer the following questions (use your markdown skills) and complete the following tasks.

2.  What type of plot would you use to display the distribution of the personal freedom scores, `pf_score`? Would this be the same type of plot to display the distribution of the political pressures and controls on media content index, `pf_expression_control`?

**A histogram would be an effective choice. A histogram allows us to see the frequency distribution of pf_score across different ranges, helping us understand its central tendency, spread, and the presence of any skewness or outliers.**

-   In the R code chunk below titled `univariable-plots`, type the R code that displays this plot for `pf_score`.
-   In the R code chunk below titled `univariable-plots`, type the R code that displays this plot for `pf_expression_control`.

```{r distribution-plots}

# Histogram for pf_score
ggplot(hfi_2016, aes(x = pf_score)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Personal Freedom Scores",
       x = "Personal Freedom Score",
       y = "Frequency") +
  theme_minimal()

# Histogram for pf_expression_control
ggplot(hfi_2016, aes(x = pf_expression_control)) +
  geom_histogram(binwidth = 1, fill = "green", color = "black") +
  labs(title = "Distribution of Political Pressures and Controls on Media Content Index",
       x = "Political Pressures and Controls on Media Content Index",
       y = "Frequency") +
  theme_minimal()
```
