---
title: "activity02"
output: html_document
date: "`r Sys.Date`"
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
```

### The data

The data we're working with is from the OpenIntro site: `https://www.openintro.org/data/csv/hfi.csv`.
Here is the "about" page: https://www.openintro.org/data/index.php?data=hfi.

In the R code chunk below titled `load-data`, you will type the code that reads in the above linked CSV file by doing the following:

- Rather than downloading this file, uploading to RStudio, then reading it in, explore how to load this file directly from the provided URL with `readr::read_csv` (`{readr}` is part of `{tidyverse}`).
- Assign this data set into a data frame named `hfi` (short for "Human Freedom Index").

```{r load data}
hfi <- read_csv("hfi.csv")
hfi
```

1. What are the dimensions of the dataset?
  What does each row represent?
  
  
```{r}
dimensions <- dim(hfi)
rows <- dimensions[1]
columns <- dimensions[2]

# Print the dimensions
cat("The dataset has", rows, "rows and", columns, "columns.\n")
```

The dataset spans a lot of years.
We are only interested in data from year 2016.
In the R code chunk below titled `hfi-2016`, type the code that does the following:

- Filter the data `hfi` data frame for year 2016, and
- Assigns the result to a data frame named `hfi_2016`.

```{r hfi-2016}
library(dplyr)

# Filter the data for year 2016 and assign it to hfi_2016
hfi_2016 <- hfi |>
  filter(year == 2016)
hfi_2016
```
### 1. Identify our research question(s)

The research question is often defined by you (or your company, boss, etc.).
Today's research question/goal is to predict a country's personal freedom score in 2016.

For this activity we want to explore the relationship between the personal freedom score, `pf_score`, and the political pressures and controls on media content index,`pf_expression_control`.
Specifically, we are going to use the political pressures and controls on media content index to predict a country's perso
